---
title: "GWR"
output: html_document
date: "2024-04-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
easypackages::packages ("sf", "sp", "spdep", "spatialreg", "GWmodel", "tmap", "mapview", "car", "RColorBrewer", 
                        "cowplot", "leafsync", "leaflet.extras2", "mapview", "tidyverse", "patchwork")
```

```{r}
project_directory <- dirname(getwd())
```

# Loading Data

## Toronto Map

```{r}
toronto_geojson_file <- file.path(project_directory, "data", "geodata", "neighbourhoods.geojson")
toronto <- st_read(toronto_geojson_file)
```

```{r}
# the AREA_NAME is the name of the neighbourhood
# we need to clean the text so it is the same as all the other files
toronto$AREA_NAME <- tolower(trimws(gsub("\\(\\d+\\)", "", toronto$AREA_NAME)))
toronto$AREA_NAME <- gsub("\\(\\d+\\)", "", toronto$AREA_NAME)
toronto$AREA_NAME <- gsub("[[:punct:]]", "", toronto$AREA_NAME)

names(toronto)[names(toronto) == "AREA_NAME"] <- "neighbourhood"


toronto$neighbourhood[which(toronto$neighbourhood == 'cabbagetownsouth stjames town')] <- 'cabbagetownsouth st james town'
toronto$neighbourhood[which(toronto$neighbourhood == 'north stjames town')] <- 'north st james town'
```

------------------------------------------------------------------------

```{r}
toronto$neighbourhood
```

## Census 2016

The census files contain 63 variables that will be used as indicators.

```{r}
#census2011_path <- file.path(project_directory, "data", "census2011.csv")
#census2011 <- read.csv(census2011_path, row.names = 1)

census2016_path <- file.path(project_directory, "data", "census2016.csv")
census2016 <- read.csv(census2016_path, row.names = 1)

#census2021_path <- file.path(project_directory, "data", "census2021.csv")
#census2021 <- read.csv(census2021_path, row.names = 1)
```

```{r}
names(census2016)
```

## Add Population Density

```{r}
toronto$area <- st_area(toronto)
```

```{r}
#census2011$population_density <- census2011$total_population/toronto$area
census2016$population_density <- census2016$total_population/toronto$area
#census2021$population_density <- census2021$total_population/toronto$area
```

## Add Toronto Geometries to Census 2016

```{r}
census2016_with_geom <- merge(census2016, toronto, by.x = "neighbourhood", by.y = "neighbourhood", all.x = TRUE)
census2016_sf <- st_as_sf(census2016_with_geom)
st_crs(census2016_sf) <- st_crs(toronto)
```

# Exploratory Spatial Data Analysis

```{r}
# change fill to see different maps
ggplot() +
  geom_sf(data = census2016_sf, aes(fill = num_crimes)) +
  scale_fill_viridis_c() +
  theme_minimal()
```

# Spatial Autocorrelation of Dependent Variable

```{r}
#Continuity
crimedata_nbq <- poly2nb(census2016_sf, queen=TRUE) 
crimedata_nbq_w <- nb2listw(crimedata_nbq, style="W", zero.policy = TRUE) #Queen’s 

coordsW <- census2016_sf%>%
  st_centroid()%>%
  st_geometry()

#KNN
#crimedata_KNN <- knearneigh(coordsW, k=4) 
#crimedata_nbq_KNN <- knn2nb(crimedata_KNN, sym=T)
#crimedata_KNN_w <- nb2listw(crimedata_nbq_KNN, style="W", zero.policy = TRUE)

plot(crimedata_nbq, st_geometry(coordsW), col="red")

mc_global <- moran.mc(census2016_sf$num_crimes, crimedata_nbq_w, 2999, alternative="greater") 

#plot the  Moran’s I
plot(mc_global)

mc_global
```

# Functions for analysis

## Linear Regression

```{r}
#calculation
linear_regression_analysis <- function(equation, data) {
    linearMod <- lm(equation, data = data)
    summary_info <- summary(linearMod)
    mc_global_OLS <- moran.mc(linearMod$residuals, crimedata_nbq_w, 2999, zero.policy = TRUE, alternative = "greater")
    vif_values <- vif(linearMod)
    
    # Save results
    linear_results <- list(model = linearMod, summary = summary_info, vif = vif_values, moran=mc_global_OLS)
    return(linear_results)
}
```

```{r}
plot_residuals <- function(linear_results){
  qtm(residuals(linear_results$linearMod))
}
```

## LISA

```{r}
# calculation
lisa_analysis <- function(data) {
    gm_crime_LISA <- localmoran(data$num_crimes, crimedata_nbq_w)
    summary_info <- summary(gm_crime_LISA)
    
    # Save results
    lisa_results <- list(gm_crime_LISA = gm_crime_LISA, summary = summary_info)
    return(lisa_results)
}
```

```{r}
# visualisation
visualise_LISA <- function(data, gm_crime_LISA) {
    sf_object <- data 
    
    sf_object$gm_crime_LISA <- gm_crime_LISA[, 1]
    p_values <- gm_crime_LISA[, 5]

    map_LISA <- tm_shape(sf_object) + 
        tm_polygons(col = "gm_crime_LISA", title = "Local Moran’s I", midpoint = 0,
                    palette = "RdYlBu", breaks = c(-1, -0.5, 0, 0.5, 1, 2))
    
    map_LISA_p <- tm_shape(sf_object) + 
        tm_polygons(col = "gm_crime_LISA", title = "p-values",
                    breaks = c(0, 0.01, 0.05, 1), palette = "Reds")

    tmap_arrange(map_LISA, map_LISA_p)
}
```

```{r}
# crime clusters 
visualize_LISA_clusters <- function(data_sf, LISA_data, significanceLevel = 0.05, colors = c('red', 'brown', 'NA', 'blue', 'cyan')) {
  # Calculate mean value of the variable under consideration
  meanVal <- mean(data_sf$num_crimes)
  
  # Create a copy of LISA data to preserve original results
  LISA_type <- LISA_data
  
  # Create a tibble object storing the LISA values into different classes of significance
  LISA_mapping <- LISA_type %>%
    tibble::as_tibble() %>%
    magrittr::set_colnames(c("Ii", "E.Ii", "Var.Ii", "Z.Ii", "Pr(z > 0)")) %>%
    dplyr::mutate(coType = dplyr::case_when(
      `Pr(z > 0)` > significanceLevel ~ "Insignificant", # Greater than significance level, thus Insignificant
      `Pr(z > 0)` <= significanceLevel & Ii >= 0 & data_sf$num_crimes >= meanVal ~ "HH", # High-High local Moran's I values
      `Pr(z > 0)` <= significanceLevel & Ii >= 0 & data_sf$num_crimes < meanVal ~ "LL", # Low-Low local Moran's I values
      `Pr(z > 0)` <= significanceLevel & Ii < 0 & data_sf$num_crimes >= meanVal ~ "HL", # High-Low local Moran's I values
      `Pr(z > 0)` <= significanceLevel & Ii < 0 & data_sf$num_crimes < meanVal ~ "LH" # Low-High local Moran's I values
    ))
  
  # Join back to the main polygon data
  data_sf$coType <- LISA_mapping$coType %>% tidyr::replace_na("Insignificant")
  
  # Map the clusters
  ggplot(data_sf) +
    geom_sf(aes(fill = coType)) +
    scale_fill_manual(values = colors, name = 'Clusters & \nOutliers') +
    labs(title = "Crime clusters") +
    theme(panel.background = element_rect(fill = 'transparent'))
}
```

## GWR

```{r}
# calculation
gwr_analysis <- function(equation, data) {
    crimedata2016_sp <- as_Spatial(data)
    abw <- bw.gwr(equation, approach = "AIC", adaptive = TRUE, kernel = "gaussian", data = crimedata2016_sp)
    a.gwr <- gwr.basic(equation, adaptive = TRUE, kernel = "gaussian", bw = abw, data = crimedata2016_sp)
    
    # Save results
    gwr_results <- list(a.gwr = a.gwr)
    return(gwr_results)
}
```

```{r}
# visualisation
extract_predictors <- function(equation) {
    predictors <- attr(terms(equation), "term.labels")
    return(predictors)
}

visualise_gwr <- function(agwr_sf, equation) {
    predictors <- extract_predictors(equation)
    
    maps <- lapply(predictors, function(predictor) {
        qtm_map <- qtm(agwr_sf, predictor)
        return(qtm_map)
    })
    
    tmap_arrange(maps, asp = 5, ncol = 1)
}
```

# Processing

## Equations

```{r}
equation1 <- num_crimes ~ unemployed + household_income + no_certificate_diploma_degree + commute_walk
equation2 <- num_crimes ~ unemployed + household_income + no_certificate_diploma_degree + commute_public_transport
equation3 <- num_crimes ~ unemployed + household_income + no_certificate_diploma_degree + commute_public_transport + population_density 
equation4 <- num_crimes ~ unemployed + household_income + no_certificate_diploma_degree + commute_public_transport + population_density + single_parents 
equation5 <- num_crimes ~ no_certificate_diploma_degree + commute_public_transport + population_density + single_parents + income_under_5000 + income_above_100k + commute_walk + non_immigrants
# List of equations
equations <- list(equation1, equation2, equation3, equation4, equation5)
```

## Calulations

```{r}
linear_results_list <- list()
lisa_results_list <- list()
gwr_results_list <- list()

# Perform analysis for each equation
for (i in seq_along(equations)) {
    eq <- equations[[i]]
    
    # Linear Regression Analysis
    linear_results_list[[i]] <- linear_regression_analysis(eq, census2016_sf)
    
    # LISA Analysis
    lisa_results_list[[i]] <- lisa_analysis(census2016_sf)
    
    # GWR Analysis
    gwr_results_list[[i]] <- gwr_analysis(eq, census2016_sf)
}
```

## Linear Regression Plots

```{r}
# equation 1
linear_results_list[[1]]
```

```{r}
# equation 2
linear_results_list[[2]]
```

```{r}
# equation 3
linear_results_list[[3]]
```

```{r}
# equation 4
linear_results_list[[4]]
```

```{r}
# equation 5
linear_results_list[[5]]
```

```{r}
census2016_sf$residuals1 <- residuals(linear_results_list[[1]]$model)
census2016_sf$residuals2 <- residuals(linear_results_list[[2]]$model)
census2016_sf$residuals3 <- residuals(linear_results_list[[3]]$model)
census2016_sf$residuals4 <- residuals(linear_results_list[[4]]$model)
census2016_sf$residuals5 <- residuals(linear_results_list[[5]]$model)
qtm(census2016_sf, "residuals1")
qtm(census2016_sf, "residuals2")
qtm(census2016_sf, "residuals3")
qtm(census2016_sf, "residuals4")
qtm(census2016_sf, "residuals5")
```

## LISA Plots

```{r}
visualise_LISA(census2016_sf, lisa_results_list[[1]]$gm_crime_LISA)
visualise_LISA(census2016_sf, lisa_results_list[[2]]$gm_crime_LISA)
visualise_LISA(census2016_sf, lisa_results_list[[3]]$gm_crime_LISA)
visualise_LISA(census2016_sf, lisa_results_list[[4]]$gm_crime_LISA)
visualise_LISA(census2016_sf, lisa_results_list[[5]]$gm_crime_LISA)
```

```{r}
visualize_LISA_clusters(census2016_sf, lisa_results_list[[1]]$gm_crime_LISA, significanceLevel = 0.05, colors = c('red', 'brown', 'NA', 'blue', 'cyan'))
visualize_LISA_clusters(census2016_sf, lisa_results_list[[2]]$gm_crime_LISA, significanceLevel = 0.05, colors = c('red', 'brown', 'NA', 'blue', 'cyan'))
visualize_LISA_clusters(census2016_sf, lisa_results_list[[3]]$gm_crime_LISA, significanceLevel = 0.05, colors = c('red', 'brown', 'NA', 'blue', 'cyan'))
visualize_LISA_clusters(census2016_sf, lisa_results_list[[4]]$gm_crime_LISA, significanceLevel = 0.05, colors = c('red', 'brown', 'NA', 'blue', 'cyan'))
visualize_LISA_clusters(census2016_sf, lisa_results_list[[5]]$gm_crime_LISA, significanceLevel = 0.05, colors = c('red', 'brown', 'NA', 'blue', 'cyan'))
```

## GWR Plots

```{r}
# Plot all GWR results for each equation
for (i in seq_along(gwr_results_list)) {
    gwr_result <- gwr_results_list[[i]]
    equation <- equations[[i]]
    
    # Extract GWR spatial data frame
    agwr_sf <- st_as_sf(gwr_result$a.gwr$SDF)
    
    # Visualize GWR result
    gwr_plot <- visualise_gwr(agwr_sf, equation)
    
    # Print GWR plot
    print(gwr_plot)
}
```
